---
title: "GAM_group_rate"
output: html_document
date: "2024-06-25"
---

## Introduction
<span> This is a full document to show the GAM model results in each subgroup for general public from 2000 to 2019. The subgroups are listed as below:
<ul>
<li>By age (age 18-34, 35-64, 65+ )  </li>
<li>By race (white, black, other races) </li>
<li>By gender (male, female) </li>
<li>By regions (West, Midwest, Southwest, Southeast, Northeast) </li>
<li>By suicide means (firearm and non-firearm) </li>
</ul>
</span>

## Loading all the dataframes
```{r}
library(mgcv)
# 1. This is the dataframe for all groups
suicide_all_groups <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_all_groups.csv')

# 2. This is the dataframe for age subgroup
suicide_age_group <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_age_group_final.csv')

# 3. This is the dataframe for gender subgroup
suicide_gender_group <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_gender_group.csv')

# 4. This is the dataframe for race subgroups
suicide_race <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_race_group.csv')

# 5. This is the dataframe for region subgroups
suicide_region <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_region.csv')

# 6. This is the dataframe for suicide means subgroups
suicide_means <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_suicide_means.csv')

```

## Change the temperature unit from F to C
```{r}
# Load the dplyr package
library(dplyr)
# Define the function to convert Fahrenheit to Celsius
fahrenheit_to_celsius <- function(fahrenheit) {
  celsius <- (fahrenheit - 32) * 5/9
  return(celsius)
}

suicide_all_groups <- suicide_all_groups %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))

suicide_age_group <- suicide_age_group %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))

suicide_gender_group <- suicide_gender_group %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))

suicide_race <- suicide_race %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))

suicide_region <- suicide_region %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))

suicide_means <- suicide_means %>%
  mutate(tMean = fahrenheit_to_celsius(tMean))
```


### Defining subgroup dataframe
```{r}
# 1. Dividing age dataframe - suicide_age_group by age_range
suicide_1534 <- suicide_age_group[suicide_age_group$age == 'age 15-34',]
suicide_3564 <- suicide_age_group[suicide_age_group$age == 'age 35-64',]
suicide_over65 <- suicide_age_group[suicide_age_group$age == '65 and older',]

# 2. Dividing gender dataframe - suicide_gender_group by gender
suicide_male <- suicide_gender_group[suicide_gender_group$sex == 'Male',]
suicide_female <- suicide_gender_group[suicide_gender_group$sex == 'Female',]

# 3. Dividing race dataframe - suicide_race by race
suicide_white <- suicide_race[suicide_race$race == 'White',]
suicide_black <- suicide_race[suicide_race$race == 'Black',]
suicide_other <- suicide_race[suicide_race$race == 'Other',]

# 4. Dividing region dataframe - suicide_region by region
suicide_west <- suicide_region[suicide_region$region == 'West',]

suicide_midwest <- suicide_region[suicide_region$region == 'Midwest',]

suicide_southwest <- suicide_region[suicide_region$region == 'Southwest',]

suicide_southeast <- suicide_region[suicide_region$region == 'Southeast',]

suicide_northeast <- suicide_region[suicide_region$region == 'Northeast',]

# 5. Dividing suicide means dataframe - suicide_means by means
suicide_firearm <- suicide_means[suicide_means$suicide_means == 'firearm_suicide',]

suicide_nonfirearm <- suicide_means[suicide_means$suicide_means == 'nonfirearm_suicide',]
```

## This the GAM model function
```{r}
model_group <- function (df) {
  # model < gam(suicide ~ s(tMean, k = 2) + s(sunshine) + s(day) + as.factor(is_holiday)) 
  # model < gam(suicide ~ s(tMean, k = 6) + s(prec, k = 6) + s(month, k = 12) + as.factor(summer_winter)) 
  model <- gam(suicide_rate ~ s(tMean), data = df)
  return(model)
}
```


### Add a vertical line at x when Y = 0 (the X value is estimated)
```{r}
add_vertical_line <- function(estimated_x) {
  x_at_y0 <- estimated_x # This value is estimated from the plot
  abline(v = x_at_y0, col = "blue", lty = 2)
  #unit <- '°C'
  text(x = x_at_y0, y = 0, labels = paste("x =", x_at_y0), pos = 3)
}
```


```{r}
plot_group <- function (model_df, group_name, df){
  plot(model_df, select=1, rug = TRUE, main = paste(group_name), 
       xlab = "Monthly Mean Temperature (°C)", ylab = "Change in Monthly Suicide Rate",
       ylim = c(-0.3, 0.3))
# add the line when y = 0
  abline(h = 0, col = "red", lty = 1)
  # Define a range of x-values to predict over
  x_range <- seq(min(df$tMean), max(df$tMean), length.out = 1000)
  x_values <- data.frame(tMean = x_range)
  base_rate <- mean(df$suicide_rate)
  # Predict y-values for the range of x-values
  predictions <- predict(model_df, newdata=x_values, se.fit=TRUE)
  y_values <- predictions$fit - base_rate
  x_at_zero <- approx(y_values, x_range, xout = 0)$y
  x_point <- round(x_at_zero, 1)
  add_vertical_line(x_point)
}

```


## Add a horizonal line at Y
```{r}
add_horizon_line <- function(estimated_y) {
  abline(h = estimated_y, col = "red", lty = 1)
}
```






```{r}
library(dplyr)

break_point <- function(df, point_list) {
  min_x <- min(df$tMean)
  max_x <- max(df$tMean)
  min_max_df <- data.frame(tMean = c(min_x, max_x))
  break_point_df <- data.frame(tMean = unlist(point_list))
  x_values <- rbind(min_max_df[1, ], break_point_df, min_max_df[2, ])
  base_rate <- mean(df$suicide_rate)
  
  model_group <- model_group(df)
  
  #[0, 3, 10, 20]
  # Predict the y values and standard errors at the specified x values
  predictions <- predict(model_group, newdata = x_values, se.fit = TRUE)

  # Extract the fitted values and standard errors
  fitted_values <- predictions$fit # suicide rate when temp at 0, 3, 10, 20 
  y_values <- fitted_values - base_rate
  standard_errors <- predictions$se.fit
  y_values_upper <- fitted_values + 1.96 * standard_errors - base_rate
  y_values_lower <- fitted_values - 1.96 * standard_errors - base_rate
  
  result_df <- x_values %>% mutate(fitted_values = fitted_values, standard_errors = standard_errors,
                                   y_values = y_values, y_values_upper = y_values_upper, 
                                   y_values_lower = y_values_lower)
  
  return(result_df)
  
}

```



```{r}
calculate_slope_details <- function(x1, y1, y1_ci_upper, y1_ci_lower, x2, y2, y2_ci_upper, y2_ci_lower, base_rate) {
  # Calculate the slope
  slope <- (y2 - y1) / (x2 - x1)
  
  # Calculate the standard error of y1 and y2 using the CI upper and lower bounds
  y1_se <- (y1_ci_upper - y1_ci_lower) / (2 * 1.96)
  y2_se <- (y2_ci_upper - y2_ci_lower) / (2 * 1.96)
  
  # Propagate error for the slope using the standard error of y1 and y2
  slope_se <- sqrt(y1_se^2 + y2_se^2) / (x2 - x1)
  
  # Calculate the 95% CI for the slope
  slope_ci_upper <- slope + (1.96 * slope_se)
  slope_ci_lower <- slope - (1.96 * slope_se)
  
  # Calculate the percentage increase based on the base rate
  percentage_increase <- (slope / base_rate) * 100
  # Calculate the standard error of the percentage increase (propagating from the slope's SE)
  percentage_increase_se <- (slope_se / base_rate) * 100
  
  # Calculate the 95% CI for the percentage increase
  percentage_increase_ci_upper <- percentage_increase + (1.96 * percentage_increase_se)
  percentage_increase_ci_lower <- percentage_increase - (1.96 * percentage_increase_se)
  
  # Return a list containing the slope, CI for slope, percentage increase, and CI for percentage increase
  return(list(
    slope = slope,
    slope_ci_upper = slope_ci_upper,
    slope_ci_lower = slope_ci_lower,
    percentage_increase = percentage_increase,
    percentage_increase_ci_upper = percentage_increase_ci_upper,
    percentage_increase_ci_lower = percentage_increase_ci_lower
  ))
}

```


```{r}
create_interval_dataframe <- function(A) {
  # Initialize an empty list to store results
  result_list <- list()
  
  # Loop over the rows of the dataframe A
  for (i in 1:(nrow(A) - 1)) {
    # Extract values for the current interval
    x1 <- A$x_values[i]
    x2 <- A$x_values[i + 1]
    y1 <- A$y_values[i]
    y2 <- A$y_values[i + 1]
    y1_ci_upper <- A$y_values_upper[i]
    y1_ci_lower <- A$y_values_lower[i]
    y2_ci_upper <- A$y_values_upper[i + 1]
    y2_ci_lower <- A$y_values_lower[i + 1]
    base_rate <- A$base_rate[i]
    group_name <- A$group[i]  # Assuming the group name is the same for consecutive rows
    
    # Calculate slope details using the provided function
    slope_details <- calculate_slope_details(
      x1, y1, y1_ci_upper, y1_ci_lower,
      x2, y2, y2_ci_upper, y2_ci_lower,
      base_rate
    )
    
    # Append the results to the list
    result_list[[i]] <- data.frame(
      group_name = group_name,
      interval_start = x1,
      interval_end = x2,
      slope = slope_details$slope,
      slope_ci_upper = slope_details$slope_ci_upper,
      slope_ci_lower = slope_details$slope_ci_lower,
      percentage_increase = slope_details$percentage_increase,
      percentage_increase_ci_upper = slope_details$percentage_increase_ci_upper,
      percentage_increase_ci_lower = slope_details$percentage_increase_ci_lower,
      base_rate = base_rate
    )
  }
  
  # Combine the list into a dataframe
  result_df <- do.call(rbind, result_list)
  
  return(result_df)
}

```


## Create a function to break the graph to two part and then use a linear function to find the coefficients in each part when the graph is a non-linear graph

```{r}
library(segmented)
convert_linear <- function(df) {
  base_rate <- mean(df$suicide_rate)
  model_df <- gam(suicide_rate ~s(tMean), data = df)
  x_range <- seq(min(df$tMean), max(df$tMean), length.out = 1000)
  x_values <- data.frame(tMean = x_range)
  predictions <- predict(model_df, newdata=x_values, se.fit=TRUE)
  change_in_suicide_rate <- predictions$fit - base_rate

  result_df <- data.frame(x_range = x_range, change_in_suicide_rate = change_in_suicide_rate)
  initial_model <- lm(change_in_suicide_rate ~ x_range, data = result_df)

  # Apply the segmented function to identify breakpoints
  segmented_model <- segmented(initial_model, seg.Z = ~x_range)

  # Extract the breakpoints
  breakpoints <- segmented_model$psi[, 2]

  print(breakpoints)
  

  # Plot the segmented model
  plot(result_df$x_range, result_df$change_in_suicide_rate, pch = 16, xlab = "Monthly Mean Temperature (°C)", ylab = "Change in Monthly Suicide Rate", ylim = c(-0.3, 0.3))
  lines(result_df$x_range, fitted(segmented_model), col = "red")
  add_vertical_line(round(breakpoints,2))

  # Add vertical lines at breakpoints
  abline(v = breakpoints, col = "blue", lty = 2)

  segments <- cut(result_df$x_range, breaks = c(-Inf, breakpoints, Inf))
  result_df$segment <- segments

  segment_models <- result_df %>%
    group_by(segment) %>%
    do(model = lm(change_in_suicide_rate ~ x_range, data = .))

  # Extract coefficients and confidence intervals
  segment_coefficients <- segment_models %>%
    summarise(
      slope = coef(model)[2],
      slope_ci_lower = confint(model)[2, 1],
      slope_ci_upper = confint(model)[2, 2]
    )

  segment_coefficients <- segment_coefficients %>%
    mutate(
      percentage_increase = (slope / base_rate) * 100,
      percentage_increase_ci_lower = (slope_ci_lower / base_rate) * 100,
      percentage_increase_ci_upper = (slope_ci_upper / base_rate) * 100,
      base_rate = base_rate
    )

  # Print the results
  print(segment_coefficients)
  return(segment_coefficients)
  }

```



### 1. Run gam model for all groups

```{r}
model_all_groups <- model_group(suicide_all_groups)
group_name1 <- 'General Public'
plot_group(model_all_groups, group_name1, suicide_all_groups)

```

```{r}
point_list_all_group <- c(NA)
result_df_all_group <- break_point(suicide_all_groups, point_list_all_group)
result_df_all_group$base_rate <- mean(suicide_all_groups$suicide_rate)
result_df_all_group$group <- 'all_group'

names(result_df_all_group)[names(result_df_all_group) == "tMean"] <- "x_values"
result_df_all_group <- na.omit(result_df_all_group)
print(result_df_all_group)

interval_df_all_group <- create_interval_dataframe(result_df_all_group)
print(interval_df_all_group)

```

### Age group - age 15-34
```{r}
model_age_1534 <- model_group(suicide_1534)
group_name1 <- 'Age 15-34'
plot_group(model_age_1534, group_name1, suicide_1534)

```

```{r}
point_list_age_1534 <- c(NA)
result_df_age_1534 <- break_point(suicide_1534, point_list_age_1534)
result_df_age_1534$base_rate <- mean(suicide_1534$suicide_rate)
result_df_age_1534$group <- 'Age 15-34'

names(result_df_age_1534)[names(result_df_age_1534) == "tMean"] <- "x_values"
result_df_age_1534 <- na.omit(result_df_age_1534)
print(result_df_age_1534)

interval_df_age_1534 <- create_interval_dataframe(result_df_age_1534)
print(interval_df_age_1534)

```


### Age group - age 35-64
```{r}
model_age_3564 <- model_group(suicide_3564)
group_name1 <- 'Age 35-64'
plot_group(model_age_3564, group_name1, suicide_3564)

```

```{r}
point_list_age_3564 <- c(NA)
result_df_age_3564 <- break_point(suicide_3564, point_list_age_3564)
result_df_age_3564$base_rate <- mean(suicide_3564$suicide_rate)
result_df_age_3564$group <- 'Age 35-64'

names(result_df_age_3564)[names(result_df_age_3564) == "tMean"] <- "x_values"
result_df_age_3564 <- na.omit(result_df_age_3564)
print(result_df_age_3564)

interval_df_age_3564 <- create_interval_dataframe(result_df_age_3564)
print(interval_df_age_3564)

```

### Age group - age 65+
```{r}
model_age_over65 <- model_group(suicide_over65)
group_name1 <- 'Age 65+'
plot_group(model_age_over65, group_name1, suicide_over65)

```

```{r}
point_list_age_over65 <- c(NA)
result_df_age_over65 <- break_point(suicide_over65, point_list_age_over65)
result_df_age_over65$base_rate <- mean(suicide_over65$suicide_rate)
result_df_age_over65$group <- 'Age 65+'

names(result_df_age_over65)[names(result_df_age_over65) == "tMean"] <- "x_values"
result_df_age_over65 <- na.omit(result_df_age_over65)
print(result_df_age_over65)

interval_df_age_over65 <- create_interval_dataframe(result_df_age_over65)
print(interval_df_age_over65)

```

## Male group
```{r}
model_male <- model_group(suicide_male)
group_name1 <- 'Male'
plot_group(model_male, group_name1, suicide_male)
```

```{r}
point_list_male <- c(NA)
result_df_male <- break_point(suicide_male, point_list_male)
result_df_male$base_rate <- mean(suicide_male$suicide_rate)
result_df_male$group <- 'male_group'

names(result_df_male)[names(result_df_male) == "tMean"] <- "x_values"
result_df_male <- na.omit(result_df_male)
print(result_df_male)

interval_df_male <- create_interval_dataframe(result_df_male)
print(interval_df_male)

```

```{r}
model_female <- model_group(suicide_female)
group_name1 <- 'Female'
plot_group(model_female, group_name1, suicide_female)

```

```{r}
point_list_female <- c(NA)
result_df_female <- break_point(suicide_female, point_list_female)
result_df_female$base_rate <- mean(suicide_female$suicide_rate)
result_df_female$group <- 'female_group'

names(result_df_female)[names(result_df_female) == "tMean"] <- "x_values"
result_df_female <- na.omit(result_df_female)
print(result_df_female)

interval_df_female <- create_interval_dataframe(result_df_female)
print(interval_df_female)

```

```{r}
model_white <- model_group(suicide_white)
group_name1 <- 'White'
plot_group(model_white, group_name1, suicide_white)

```

```{r}
point_list_white <- c(NA)
result_df_white <- break_point(suicide_white, point_list_white)
result_df_white$base_rate <- mean(suicide_white$suicide_rate)
result_df_white$group <- 'White_group'

names(result_df_white)[names(result_df_white) == "tMean"] <- "x_values"
result_df_white <- na.omit(result_df_white)
print(result_df_white)

interval_df_white <- create_interval_dataframe(result_df_white)
print(interval_df_white)

```




```{r}
model_black <- model_group(suicide_black)
group_name1 <- 'Black'
plot_group(model_black, group_name1, suicide_black)

```

```{r}
point_list_black <- c(NA)
result_df_black <- break_point(suicide_black, point_list_black)
result_df_black$base_rate <- mean(suicide_black$suicide_rate)
result_df_black$group <- 'Black_group'

names(result_df_black)[names(result_df_black) == "tMean"] <- "x_values"
result_df_black <- na.omit(result_df_black)
print(result_df_black)

interval_df_black <- create_interval_dataframe(result_df_black)
print(interval_df_black)

```


```{r}
model_other <- model_group(suicide_other)
group_name1 <- 'Other'
plot_group(model_other, group_name1, suicide_other)

```

```{r}
interval_df_other <- convert_linear(suicide_other)

```

```{r}
interval_df_other$group_name <- 'Other'
interval_breakpoint <- 11.55
interval_start <- c(min(suicide_other$tMean), interval_breakpoint)
interval_end <- c(interval_breakpoint, max(suicide_other$tMean))
interval_df_other$interval_start <- interval_start
interval_df_other$interval_end <- interval_end
interval_df_other <- interval_df_other[, c('group_name', 'interval_start', 'interval_end', 'slope', 'slope_ci_upper', 'slope_ci_lower', 'percentage_increase', 'percentage_increase_ci_upper', 'percentage_increase_ci_lower', 'base_rate')]
interval_df_other

```



```{r}
point_list_other <- c(NA)
result_df_other <- break_point(suicide_other, point_list_other)
result_df_other$base_rate <- mean(suicide_other$suicide_rate)
result_df_other$group <- 'Other_races'

names(result_df_other)[names(result_df_other) == "tMean"] <- "x_values"
result_df_other <- na.omit(result_df_other)
print(result_df_other)

```

```{r}
model_firearm <- model_group(suicide_firearm)
group_name1 <- 'Firearm'
plot_group(model_firearm, group_name1, suicide_firearm)

```

```{r}
point_list_firearm <- c(NA)
result_df_firearm <- break_point(suicide_firearm, point_list_firearm)
result_df_firearm$base_rate <- mean(suicide_firearm$suicide_rate)
result_df_firearm$group <- 'Firearm'

names(result_df_firearm)[names(result_df_firearm) == "tMean"] <- "x_values"
result_df_firearm <- na.omit(result_df_firearm)
print(result_df_firearm)

interval_df_firearm <- create_interval_dataframe(result_df_firearm)
print(interval_df_firearm)

```

```{r}
model_nonfirearm <- model_group(suicide_nonfirearm)
group_name1 <- 'Non-firearm'
plot_group(model_nonfirearm, group_name1, suicide_nonfirearm)

```

```{r}
point_list_nonfirearm <- c(NA)
result_df_nonfirearm <- break_point(suicide_nonfirearm, point_list_nonfirearm)
result_df_nonfirearm$base_rate <- mean(suicide_nonfirearm$suicide_rate)
result_df_nonfirearm$group <- 'Non-Firearm'

names(result_df_nonfirearm)[names(result_df_nonfirearm) == "tMean"] <- "x_values"
result_df_nonfirearm <- na.omit(result_df_nonfirearm)
print(result_df_nonfirearm)

interval_df_nonfirearm <- create_interval_dataframe(result_df_nonfirearm)
print(interval_df_nonfirearm)

```


```{r}
model_west <- model_group(suicide_west)
group_name1 <- 'West'
plot_group(model_west, group_name1, suicide_west)

```

```{r}
point_list_west <- c(NA)
result_df_west <- break_point(suicide_west, point_list_west)
result_df_west$base_rate <- mean(suicide_west$suicide_rate)
result_df_west$group <- 'West'

names(result_df_west)[names(result_df_west) == "tMean"] <- "x_values"
result_df_west <- na.omit(result_df_west)
print(result_df_west)

interval_df_west <- create_interval_dataframe(result_df_west)
print(interval_df_west)

```


```{r}
model_midwest <- model_group(suicide_midwest)
group_name1 <- 'Midwest'
plot_group(model_midwest, group_name1, suicide_midwest)

```

```{r}
point_list_midwest <- c(NA)
result_df_midwest <- break_point(suicide_midwest, point_list_midwest)
result_df_midwest$base_rate <- mean(suicide_midwest$suicide_rate)
result_df_midwest$group <- 'Midwest'

names(result_df_midwest)[names(result_df_midwest) == "tMean"] <- "x_values"
result_df_midwest <- na.omit(result_df_midwest)
print(result_df_midwest)

interval_df_midwest <- create_interval_dataframe(result_df_midwest)
print(interval_df_midwest)

```


```{r}
model_southwest <- model_group(suicide_southwest)
group_name1 <- 'Southwest'
plot_group(model_southwest, group_name1, suicide_southwest)

```

```{r}
point_list_southwest <- c(NA)
result_df_southwest <- break_point(suicide_southwest, point_list_southwest)
result_df_southwest$base_rate <- mean(suicide_southwest$suicide_rate)
result_df_southwest$group <- 'Southwest'

names(result_df_southwest)[names(result_df_southwest) == "tMean"] <- "x_values"
result_df_southwest <- na.omit(result_df_southwest)
print(result_df_southwest)

interval_df_southwest <- create_interval_dataframe(result_df_southwest)
print(interval_df_southwest)

```

```{r}
model_southeast <- model_group(suicide_southeast)
group_name1 <- 'Southeast'
plot_group(model_southeast, group_name1, suicide_southeast)

```

```{r}
point_list_southeast <- c(NA)
result_df_southeast <- break_point(suicide_southeast, point_list_southeast)
result_df_southeast$base_rate <- mean(suicide_southeast$suicide_rate)
result_df_southeast$group <- 'Southeast'

names(result_df_southeast)[names(result_df_southeast) == "tMean"] <- "x_values"
result_df_southeast <- na.omit(result_df_southeast)
print(result_df_southeast)

interval_df_southeast <- create_interval_dataframe(result_df_southeast)
print(interval_df_southeast)

```

```{r}
model_northeast <- model_group(suicide_northeast)
group_name1 <- 'Northeast'
plot_group(model_northeast, group_name1, suicide_northeast)

```

```{r}
point_list_northeast <- c(NA)
result_df_northeast <- break_point(suicide_northeast, point_list_northeast)
result_df_northeast$base_rate <- mean(suicide_northeast$suicide_rate)
result_df_northeast$group <- 'Northeast'

names(result_df_northeast)[names(result_df_northeast) == "tMean"] <- "x_values"
result_df_northeast <- na.omit(result_df_northeast)
print(result_df_northeast)

interval_df_northeast <- create_interval_dataframe(result_df_northeast)
print(interval_df_northeast)

```


```{r}
df_list_point <- list(result_df_all_group, result_df_age_1534, result_df_age_3564, 
                      result_df_age_over65, result_df_male, result_df_female, 
                      result_df_white, result_df_black, result_df_other,
                      result_df_firearm, result_df_nonfirearm, result_df_west,
                      result_df_midwest, result_df_southwest, result_df_southeast,
                      result_df_northeast)

# Use do.call with rbind to combine the list of data frames
point_combined_df <- do.call(rbind, df_list_point)

print(point_combined_df)

```

```{r}
df_list_interval <- list(interval_df_all_group, interval_df_age_1534, interval_df_age_3564,
                        interval_df_age_over65, interval_df_male, interval_df_female, 
                        interval_df_white, interval_df_black, 
                        interval_df_other, interval_df_firearm, interval_df_nonfirearm, 
                        interval_df_west, interval_df_midwest, interval_df_southwest,
                        interval_df_southeast, interval_df_northeast)

# Use do.call with rbind to combine the list of data frames
interval_combined_df <- do.call(rbind, df_list_interval)

print(interval_combined_df)

```

```{r}
# Save the data frame to a CSV file
write.csv(point_combined_df, file = '~/Desktop/GAM_final/GAM_group_breakpoints.csv', row.names = FALSE)
write.csv(interval_combined_df, file = '~/Desktop/GAM_final/GAM_group_percent_increase.csv', row.names = FALSE)
```


## Check year range 1968 to 2004
```{r}
# 2. This is the dataframe for age subgroup
suicide_6804 <-read.csv('~/Desktop/GAM_final/GAM_monthly_dataframes/monthly_GAM_all_groups_1968_2004.csv')

#suicide_6804 <- suicide_6804 %>%
#  mutate(tMean = fahrenheit_to_celsius(tMean))

```


```{r}
model_6804 <- model_group(suicide_6804)
group_name1 <- '1968 to 2004'
plot_group(model_6804, group_name1, suicide_6804)

```

```{r}
convert_linear(suicide_6804)
```



```{r}
point_list_6804 <- c(-2)
result_df_6804 <- break_point(suicide_6804, point_list_6804)
result_df_6804$base_rate <- mean(suicide_6804$suicide_rate)
result_df_6804$group <- '1968 to 2004'

names(result_df_6804)[names(result_df_6804) == "tMean"] <- "x_values"
result_df_6804 <- na.omit(result_df_6804)
print(result_df_6804)

interval_df_6804 <- create_interval_dataframe(result_df_6804)
print(interval_df_6804)

```


